---
title: "Yu Group software series: R"
subtitle: "R dev | simChef + dgpoix | vdocs"
author: "Tiffany Tang and James Duncan"
date: "2022/04/27"
output:
  xaringan::moon_reader:
    self_contained: true
    lib_dir: libs # path to local remark.js library
    css: xaringan-themer.css
    nature:
      highlightStyle: googlecode # solarized-dark
      countIncrementalSlides: false
      navigation:
        scroll: false
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
xaringanthemer::style_mono_accent(
  base_color = "#00b0da",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
xaringanExtra::use_fit_screen()
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}

.remark-code, .remark-inline-code  { font-family: 'Ubuntu Mono'; }
a, a > .remark-inline-code { color: rgb(249, 38, 114); }
.remark-slide-scaler { overflow-y: auto; }
```

# testthat

Testing is an important part of the data science workflow!
`r emo::ji("paperclips")`[`testthat`](https://testthat.r-lib.org/)  makes it simple.

```{r, eval=FALSE}
hitchhiker <- function(x) {
  return(x + 1)
}

testthat::expect_equal(hitchhiker(41), 42)
testthat::expect_equal(hitchhiker(42), 42)

## Error: hitchhiker(42) not equal to 42.
## 1/1 mismatches
## [1] 43 - 42 == 1
```
Other useful `testthat` verbs:
* `expect_identical` -- test if objects are the same
* `expect_true`, `expect_false`
* `expect_lt`, `expect_gt`
* `expect_error`, `expect_warning`
* `expect_snapshot` -- saves stdout and later compares with this gold standard

---

# testthat

`r emo::ji("file_folder")` `testthat` tests should be in a directory called `tests/testthat/` next to your
project's `R/` directory. (`usethis::use_testthat` can set that up for you.) 

`r emo::ji("page_facing_up")` `testthat` test files should be named like `test-*.R`, e.g. `test-xyz.R`. Again,
there's a `usethis` function to create this: `usethis::use_test("xyz")`.

`r emo::ji("+1")` create one `test-*.R` file for every `*.R` file in `R/`, e.g.,

Here's an example from `simChef`:
```{r, eval=FALSE}
test_that("fit_experiment works properly with future::multicore", {

  # multicore plan isn't supported on windows
  skip_on_os("windows")

  # fit_experiment_fixture() returns sequential and parallel simulation results
  results <- fit_experiment_fixture(future::multicore)

  expect_identical(results[[1]], results[[2]])
})
```

---

# withr

`r emo::ji("paperclips")`[`withr`](https://withr.r-lib.org/index.html) is another helpful package for
writing testing code and much more.

* Prevents unintended side effects (aka, hard to squash `r emo::ji("mosquito")`!): `with_environment()`
* Cleans up temporary state: `with_tempfile()`
* Helps with reproducibility: `with_seed()`
* Can run arbitrary code when exiting a frame in the call stack: `defer()`, `defer_parent()`

---

# withr in simChef

We use `withr` to temporarily set a `future` plan and reset it when the code
inside of `with_plan` finishes.

```{r, eval=FALSE}
set_plan <- function(plan, ...) {
  # get the original plan so we can reset at the end
  old_plan <- future::plan()
  future::plan(plan, ...)
  return(old_plan)
}

reset_plan <- function(plan) {
  future::plan(plan)
}

with_plan <- withr::with_(set_plan, reset_plan)

fit_experiment_fixture <- function(plan) {
  ... # setup code

  parallel <- withr::with_seed( # set the random seed
    seed,
    with_plan( # set the plan
      plan,
      fit_experiment(experiment, n_reps = 10) # get results in parallel
    )
  )
  return(list(parallel, sequential))
}
```

---

# GitHub Actions

 `r emo::ji("paperclips")`[GitHub Actions](https://docs.github.com/en/actions): automated scripts that run
on GitHub's servers when you make changes to a repo.

They help automate:

* testing
* code "linting" (syntax / style checking)
* (documentation) website deployment
* many other repetitive tasks that run when code changes

Here's the `usethis` command: `usethis::use_github_actions("check-standard")`.

---

# Example: Automated grading of STAT 215A lab reports

`r emo::ji("paperclips")`[auto215a](https://github.com/Yu-Group/auto215a/) (private repo)



---

# `Yu-Group` Actions

GitHub Action are a key part of the group's project / software development
workflow, used by:

* `simChef`
* `vdocs`
* `vflow`
* `imodels`
* Many more projects and hopefully yours too!

---

# r-lib GitHub actions

Many of the core R packages we've mentioned are maintained in the
`r emo::ji("paperclips")`[`r-lib`](https://github.com/r-lib) GitHub organization, including:
* `devtools`
* `roxygen2`
* `testthat`
* `pkgdown`
* `usethis`

`r-lib` also maintains a very useful repo with many GitHub Actions scripts that
you can use in your workflows: `r emo::ji("paperclips")`[r-lib/actions](https://github.com/r-lib/actions).

---

# xaringan

We created these slides in `Rmarkdown` via the package `r
emo::ji("paperclips")`[`xaringan`](https://bookdown.org/yihui/rmarkdown/xaringan.html).

How these slides were created:

```{r, eval=FALSE}
install.packages("xaringan")

# download deps for offline viewing of slides
xaringan::summon_remark()

# create an html version of the slides
rmarkdown::render("slides.Rmd")

# push the slides to GitHub pages
# see `?gh_token_help` for info on creating / registering a PAT
# TODO: the repo needs to be public for this to work
usethis::use_github_pages(path = "slides")
```
* `r emo::ji("paperclips")`[xaringanthemer](https://pkg.garrickadenbuie.com/xaringanthemer)
* `r emo::ji("paperclips")`[xaringanExtra](https://pkg.garrickadenbuie.com/xaringanExtra)

---

# Additional resources

* `r emo::ji("paperclips")`["Advanced R" (2nd edition)](https://adv-r.hadley.nz)
* `r emo::ji("paperclips")`["R Packages" (2nd edition)](https://r-pkgs.org/)
* `r emo::ji("paperclips")`["bookdown: Authoring Books and Technical Documents with R Markdown"](https://bookdown.org/yihui/bookdown/)
